Debugging by {\huge\bf\textit{lastChange}}

### 2011-03-10 papers/fse2011/sig-alternate.tex

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{subfigure} %[tight,normalsize]
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{textcomp}
\usepackage{listings}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{cite} %after hyperref

@@

A common strategy for locating defects starts from bug symptoms and
works backwards, moving from a point in the program execution where a
value appears to be incorrect back to the point were that value was
set.  Two conventional approaches, breakpoint-based and
log-based debugging, require tedious steps of
selecting data to be collected, collecting the data, then analyzing
the results.

@@

Our contribution in this paper is the technique \textit{lastChange},
which locates the last place a value has changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The technique builds on existing breakpoint debugger
technology  and it does not require a special environment to
create identical, instruction by instruction, re-executions. We demonstrate the feasibility of the approach with 
\textit{Querypoint}, an implementation extending Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.
The \textit{lastChange} algorigthm provides information on important program
values during the program execution without voluminous logs and without tedious
 insertion and removal of breakpoints. We believe other queries over the
 running program can be formulated to generalize this technique.

@@

By browsing through the code or other means\cite{Barton},
 the developer determines that the value displayed on the 
button is set at line 22. Since the displayed value is incorrect we know 
the bug occured before we hit this line.
To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, the execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

@@

\includegraphics[width=.48\textwidth]{5-example-points.png}
\caption{The examined points before locating the defect. The arrow represents the logical forward progress of the program. Three actual executions are superimposed on this arrow. All three stop at the reproduction point indicated by circle 1. After the first execution, the developer asks for lastChange as described in section \ref{sec:introExample}, yielding information  indicated by circle 2. After the second execution, another lastChange query causes a third execution, yielding information indicated by circle 3.}
\label{fig:example-points}
\end{figure}

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer runs the \textit{lastChange} command
on \texttt{myCondition.value} at this point. The debugger re-executes the
program and breaks again on line 22, analyzes its queries and shows the developer line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value}. (If the developer asks for \textit{lastChange} on \texttt{oldValue}, 
the debugger can notify the developer that this variable is never assigned a value).
 Now it is clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once the execution reaches line 25 (Figure~\ref{fig:example4}).

@@

\subfigure[A screen shot of the Firebug debugger while running the example code from Fig.~\ref{fig:js-code}. The Script
  panel is selected; it gives access to
  all loaded source files and allows breakpoints to be set on lines. In this
  figure, the execution is paused at line 22 by a regular
  breakpoint. The Watch panel on the right shows the program state at
  the paused
  point. Developer can query \textit{lastChange} on \texttt{myObject.myProperty} by right-clicking on the value of \texttt{myProperty}. ]{\label{fig:example1}\includegraphics[width=1.0\textwidth, height=.17\textheight
  ]{1-bp22.jpg}}

@@

Notice that in our example, \textit{lastChange} combines some aspects
of breakpoint and of log-based debugging. Like breakpoint debugging,
the developer re-executes a live runtime without changing the source
and without a special execution environment beyond the debugger. The
state of the program memory and the call stack are available at each
lastChange point. Like log-based debugging, the program state and the
call stack are recorded during program execution. We can't halt the
program at \textit{lastChange} because we don't know which point is the last
one until we return to the original breakpoint. In section 5 we
discuss cases were it is possible to pause at lines of \textit{lastChange}.

@@

\begin{figure}[htp]
\centering 
\includegraphics [width=.48\textwidth] {6-foo-changes1.png} %[width=.48\textwidth][scale=0.75]
\caption{A hypothetical list of  change events for a property \texttt{foo}. 
Each change event adds a column with the id of the object changed, the call stack,
and some program state such as local variable values. 
At the reproduction point  we determine which id  corresponds to 
object \texttt{bar} and read out column 4,  the last
  change of \texttt{bar.foo}. Column 3 is also a change of
  the object we want to study, id 1010, but it is not the last change;
  Column 5 is also a change of a property \texttt{foo} but it is not for
  the object we are interested in.}
\label{fig:foo-changes1}
\end{figure}

@@

\begin{figure}[htp]
\centering 
\includegraphics[width=.40\textwidth]{7-foo-changes2.png}
\caption{A hypothetical list of variable \texttt{foo} change events. Each column of the list indicates a change event; for each
change event the scope id returns by \texttt{scopeId()} is recorded along with call stack and program state information. 
   The last column having the scope id \texttt{foo} at the reproduction point indicates 
  the last change.}
\label{fig:foo-changes2}
\end{figure}

@@

\begin{figure}[htp]
\centering 
\includegraphics[width=.48\textwidth]{8-lastchange-lastchange.png}
\caption{The list of change events stored for locating point B, the
  \textit{lastChange} of \texttt{bar.x} at the reproduction point, and
  point C, the \textit{lastChange} of \texttt{baz.y} at point B.}
\label{fig:lastchange-lastchange}
\end{figure}

@@

% Cut to reduce size
%There are cases where event indexes fail, but a developer can use knowledge gained from \textit{lastChange}
%to formulate a conditional breakpoint.
%Consider the example in Figure~\ref{fig:counter-example}. It shows a function which processes an
%array. The array contains numbers except one item which is
%\texttt{undefined} and it causes a bug in line 30. There is a call to
%\texttt{randomPermutation} function in line 11 which randomly
%permutes the array item. So at every execution the
%\texttt{undefined} item will be in a new place. Calling
%\textit{lastChange} on \texttt{x} at the place the bug happens, gives a
%point which shows line 14. Although this point exists at every
%re-execution but it can not be identified by an event index. A conditional breakpoint on 
%the value of the item would succeed and the developer can reason from the 
\%textit{lastChange} results to create the condition.

>>>>

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subfigure} %[tight,normalsize]
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{textcomp}
\usepackage{listings}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{cite} %after hyperref

@@

A common strategy for locating defects starts from bug symptoms and
works backwards, moving from a point in the program execution where a
value appears to be incorrect back to the point where that value was
set.  Two conventional approaches, breakpoint-based and
log-based debugging, require tedious steps of
selecting data to be collected, collecting the data, then analyzing
the results.

@@

Our contribution in this paper is the technique \textit{lastChange},
which locates the last place a value has changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The technique builds on existing breakpoint debugger
technology  and it does not require a special environment to
create identical, instruction by instruction, re-executions. We demonstrate the feasibility of the approach with 
\textit{Querypoint}, an implementation extending Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.
The \textit{lastChange} algorithm provides information on important program
values during the program execution without voluminous logs and without tedious
 insertion and removal of breakpoints. We believe other queries over the
 running program can be formulated to generalize this technique.

@@

By browsing through the code or other means\cite{Barton},
 the developer determines that the value displayed on the 
button is set at line 22. Since the displayed value is incorrect we know 
the bug occurred before we hit this line.
To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, the execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is \texttt{0}. We expected this value to be \texttt{2}.

@@

\includegraphics [width=.48\textwidth]{5-example-points.pdf} 
\caption{The examined points before locating the defect. The arrow represents the logical forward progress of the program. Three actual executions are superimposed on this arrow. All three stop at the reproduction point indicated by circle 1. After the first execution, the developer asks for lastChange as described in section \ref{sec:introExample}, yielding information  indicated by circle 2. After the second execution, another lastChange query causes a third execution, yielding information indicated by circle 3.}
\label{fig:example-points}
\end{figure}

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer runs the \textit{lastChange} command
on \texttt{myCondition.value} at this point. The debugger re-executes the
program and breaks again on line 22, analyzes its queries and shows the developer line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value}. If the developer asks for \textit{lastChange} on \texttt{oldValue}, 
the debugger can notify the developer that this variable is never assigned a value.
 Now it is clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once the execution reaches line 25 (Figure~\ref{fig:example4}).

@@

\subfigure[A screen shot of the Firebug debugger while running the example code from Fig.~\ref{fig:js-code}. The Script
  panel is selected; it gives access to
  all loaded source files and allows breakpoints to be set on lines. In this
  figure, the execution is paused at line 22 by a regular
  breakpoint. The Watch panel on the right shows the program state at
  the paused
  point. Developer can query \textit{lastChange} on \texttt{myObject.myProperty} by right-clicking on the value of \texttt{myProperty}. ]{\label{fig:example1}\includegraphics[width=1.0\textwidth, %height=.17\textheight
  ]{1-bp22.jpg}}

@@

Notice that in our example, \textit{lastChange} combines some aspects
of breakpoint and of log-based debugging. Like breakpoint debugging,
the developer re-executes a live runtime without changing the source
and without a special execution environment beyond the debugger. The
state of the program memory and the call stack are available at each
lastChange point. Like log-based debugging, the program state and the
call stack are recorded during program execution. We can't halt the
program at \textit{lastChange} because we don't know which point is the last
one until we return to the original breakpoint. In section 5 we
discuss cases where it is possible to pause at lines of \textit{lastChange}.

@@

\begin{figure}[htp]
\centering 
\includegraphics [width=.48\textwidth] {6-foo-changes1.pdf} %[width=.48\textwidth][scale=0.75]
\caption{A hypothetical list of  change events for a property \texttt{foo}. 
Each change event adds a column with the id of the object changed, the call stack,
and some program state such as local variable values. 
At the reproduction point  we determine which id  corresponds to 
object \texttt{bar} and read out column 4,  the last
  change of \texttt{bar.foo}. Column 3 is also a change of
  the object we want to study, id 1010, but it is not the last change;
  Column 5 is also a change of a property \texttt{foo} but it is not for
  the object we are interested in.}
\label{fig:foo-changes1}
\end{figure}

@@

\begin{figure}[htp]
\centering 
\includegraphics[width=.40\textwidth]{7-foo-changes2.pdf}
\caption{A hypothetical list of variable \texttt{foo} change events. Each column of the list indicates a change event; for each
change event the scope id returns by \texttt{scopeId()} is recorded along with call stack and program state information. 
   The last column having the scope id \texttt{foo} at the reproduction point indicates 
  the last change.}
\label{fig:foo-changes2}
\end{figure}

@@

\begin{figure}[htp]
\centering 
\includegraphics[width=.48\textwidth]{8-lastchange-lastchange.pdf}
\caption{The list of change events stored for locating point B, the
  \textit{lastChange} of \texttt{bar.x} at the reproduction point, and
  point C, the \textit{lastChange} of \texttt{baz.y} at point B.}
\label{fig:lastchange-lastchange}
\end{figure}

@@

% Cut to reduce size
%There are cases where event indexes fail, but a developer can use knowledge gained from \textit{lastChange}
%to formulate a conditional breakpoint.
%Consider the example in Figure~\ref{fig:counter-example}. It shows a function which processes an
%array. The array contains numbers except one item which is
%\texttt{undefined} and it causes a bug in line 30. There is a call to
%\texttt{randomPermutation} function in line 11 which randomly
%permutes the array item. So at every execution the
%\texttt{undefined} item will be in a new place. Calling
%\textit{lastChange} on \texttt{x} at the place the bug happens, gives a
%point which shows line 14. Although this point exists at every
%re-execution but it can not be identified by an event index. A conditional breakpoint on 
%the value of the item would succeed and the developer can reason from the 
%textit{lastChange} results to create the condition.
